Some thought experiments on the applicability of Granger causality and Directed Information in inferring direction of information flows

### 2015-07-06 paper.tex
Add section on GC used in neuroscience.

\documentclass[conference]{IEEEtran}

@@

\subsection{Motivation}
\label{subsec:motivation}

@@

\begin{enumerate}
	\item We take Brovelli et.\ al.\ and Ding et.\ al.\ as the basis / canonical reference, as this appears to be generally accepted by the community (Cite review papers that cite these two papers; cite Anil Seth's MVGC package paper).
	\item Explain Granger causality
	\item Explain the exact method used - including trials, pre-processing steps, etc.\ - as described in the Bressler paper.
	\item Granger's original analysis does not compare GC on forward and reverse links, however, this appears to be the commonly done in practice (cite Brovelli and review papers if possible)
	\item State our problem statement again - we take issue not with any of these steps, but with the interpretation that if greater Granger causal influences are seen to be from X to Y, then information must be flowing from X to Y in the brain.
\end{enumerate}

\subsection{Our counter-examples and main result}

\subsection{Survey of previous criticisms of Granger Causality}

@@

>>>>

\documentclass[conference]{IEEEtran}
%\documentclass[conference,draftcls]{IEEEtran}

@@

\subsection{Motivation}
\label{sec:motivation}

@@

We take~\cite{Brovelli2004BetaOscillations} and~\cite{Ding1999ShortWindow} to be the canonical references for the application of Granger causality to estimate causal influences, as they seem to be widely accepted by the community (\cite{Bressler2011WienerGranger},{\color{red} \cite{Roebroeck2005MappingDirected}},~\cite{Ding2006GrangerCausality}, \cite{Barnett2014MVGC}). Here, we briefly describe how Granger Causal influence is quantified, and how it is computed in these works.

Granger causality, as originally described by Granger~\cite{Granger1969InvestigatingCausal}, measures the level of influence that one process $\{X\}$ has on another process $\{Y\}$. The analysis compares the \textit{error in predicting} the $\{Y\}$ process based on (i) simply the past of $\{Y\}$, and (ii) based on the past of both $\{X\}$ and $\{Y\}$. The Granger causality metric is the ratio of these errors. Many variations of Granger causality have also been considered, including a generalization --- the Directed Information (see~\cite{MasseyCausality,quinn,weissman}) --- an information-theoretic quantity denoted by $I(\m{X}\to\m{Y})$.

Granger causality computation, in general, requires that the processes in consideration be stationary. In neuroscientific experiments, this is usually not the case, as the express objective of the experiment is to understand brain function while the brain is subjected to a task. In order to circumvent this problem, data analysis techniques either assume stationarity~\cite{Brovelli2004BetaOscillations}, or work with time-windows short enough that the processes are in fact stationary within those windows~\cite{Ding2006GrangerCausality}. In the latter case, Granger causal influences are computed for each window individually, in order to produce a dynamic description of brain connectivity as the brain carries out the said task.

In order to determine the direction of causal influence, the Granger Causality metric (ratio of residual variances) from $\{X\}$ to $\{Y\}$ is \emph{compared} to that from $\{Y\}$ to $\{X\}$. We note here that Granger's original analysis does not compare this metric on forward and reverse links, however, this appears to be commonly done in practice. {\color{red} TODO: Cite Tsachy's paper here.}

While we accept that Granger Causal influence can be accurately computed (provided measurements are taken suitably~\ref{sec:previous-criticisms-of-gc}) and can be useful in many situations~\ref{sec:discussion}, interpreting the direction of Granger Causal influence as the direction of information flow can be erroneous, as we demonstrate in this paper.

\subsection{A short survey of previous criticisms of Granger Causality}

Several objections have been leveled against Granger Causality in the past. We give, here, a short overview of these and describe why our objection is novel, and possibly more fundamental in nature, at least in the context of its usage in neuroscience.

@@

\subsection{Our counter-examples and our main result}
